DeepSphere: a graph-based spherical CNN

### 2019-09-05 main.tex
cleanup and first pass

%\title{Which spherical CNN should you use?\\ DeepSphere V2}
\title{DeepSphere: an efficient graph-based spherical CNN}
%\title{DeepSphere, an efficient (& equivariant) graph-based spherical CNN}

@@

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

@@

contributions:\\
* theory => convergence (better graph)\\
* experiments on relevant problems (not spherical MNIST) => check the desiderata\\
  * surprising: anisotropy doesn't seem useful\\

@@

@@

Desiderata:\\
* respect the domain\\
  * rotational equivariance\\
  * deformation (e.g., on the icosahedron for gauge)\\
* powerful / generality => isotropic vs anisotropic\\
* scale: computational cost \& memory usage (feature maps)\\
* flexible (samplings, irregular), simplicity (implementation)\\

\begin{table}[h!]
    \centering
    \begin{tabular}{l|c|c|c|c}
         & Scale & Generality& equivariance & Flexibility \\
         \hline
        Cohen & & & &  \\
         \hline
        Gauge & & & &  \\
         \hline
        Esteves & & & &  \\
         \hline
        Jiang & & & &  \\
         \hline
        2D CNN (cubed sphere) & & & &  \\
         \hline
        Ours & & & &  \\
    \end{tabular}
    \caption{Caption}
    \label{tab:my_label}
\end{table}

On the one side, we have Cohen which is computationally very expensive but perfectly equivariant, on the other we have cube sphere which computationl very good but not equivariant at all. Other methods are tradedoff between these two. What is a good tradeoff? We think that DeepSphere is

Present the different spherical CNNs with pro and cons => mostly scale vs generality\\
* 2D CNN => doesn't respect geometry and equivariance\\
* Cohen spherical CNN => most general, but doesn't scale\\
* Cohen gauge => fixes scale, at the price of deformation (sphere => icosahedron)\\
* Esteves => ? \\
* Jiang => need a global coordinate system (ok for planets, not projections like cosmo)\\

@@

* HEALPix: improvement from DeepSphere v1 \cite{perraudin2019deepsphere}
* equiangular: Renata \& Pascal

@@

\subsection{HEALPix: convergence theorem}
Given a sampling $x_0, ..., x_{n-1}$ define $\sigma_i$ to be the patch of the surface of the sphere corresponding to the $i$th point of the sampling, define $A_i$ to be its corresponding area and $d_i$ to be the radius of the smallest ball in $\mathbb R^3$ containing the $i$th patch. Define $d^{(n)} := \max_{i=0, ..., n}d_i$ and $A^{(n)}=\max_{i=0, ..., n}A_i$.\\
	\begin{theorem}
		For a sampling $\mathcal P = \{x_i\in\mathbb S^2\}_{i=0}^{n-1}$ of the sphere that is equi area and such that $d^{(n)})\leq \frac{C}{\sqrt{n}}$, for all $f: \mathbb S^2 \rightarrow \mathbb R$ Lipschitz with respect to the euclidean distance in $\mathbb R^3$, for all $y\in\mathbb S^2$, there exists a sequence $t_n = n^\beta$ such that the rescaled Heat Kernel Graph Laplacian operator $\frac{|\mathbb S^2|}{4\pi t_n}L^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\triangle_{\mathbb S^2}$  for $n\to\infty$:
		$$ \lim_{n\to\infty}\frac{|\mathbb S^2|}{4\pi t_n} L_n^{t_n}f(y) =  \triangle_{\mathbb S^2}f(y).$$
		\label{theo:pointwise convergence in the healpix case}
	\end{theorem}

\subsection{Equiangular}
Khasanova et al. \cite{Frossard2017GraphBasedCO} designed a discrete Laplacian that is explicitly intended to work on the sphere with the equiangular sampling. They consider the set $\mathcal G$ of all the possible graphs where each node is connected only to four of its nearest neighbours (North, Sud, West, East) and propose a weighting scheme $w_{ij}$ that minimizes the difference in the response to the polynomial spectral filter $\mathcal F = \mathbf L$ evaluated on images of the same object seen at different latitudes. In other words, they solve the minimization problem

\begin{equation}\label{eq:minimization frossard}
	\min_{W\in\mathcal G} \left|\mathcal{F}\left(\mathbf{y}\left(v_{ e}\right)\right)-\mathcal{F}\left(\mathbf{y}\left(v_{ i}\right)\right)\right|
\end{equation}
for the adjacency matrix $W$, where $\mathbf y(v_i)$ is the image of the object on the sphere centered on the vertex $v_i$, and $\mathcal F (\mathbf y(v_e))$ is the response of the filter at the vertex $v_e$ that lies at the same longitude of the vertex $v_i$ but on the equator. In their work they prove that the optimal weights solving the minimization problem (\ref{eq:minimization frossard}) are given by weights $w_{ij}$ inversely proportional to the Euclidean distance between vertices:
\begin{equation}\label{eq:frossard weights}
	w_{ij} = \frac{1}{\norm{x_i-x_j}}
\end{equation}
This construction is interesting since it is adapted to the equiangular sampling, and leads to a very sparse graph with only 4 neighbors per vertex. Furthermore, to obtain the weights (\ref{eq:frossard weights}) every calculation was done in the \textit{spatial domain}, without any consideration about the spectral interpretation of the filter. In order to compare it to the HKGL we show the equivariance error by spherical harmonic degree in figure \ref{fig:Equivariance error of the Frossard-Khasanove graph}. It can be appreciated how this construction performs a little worse that the full HKGL for low bandwidth samplings, but converges faster than the \textit{full} HKGL, and its spectrum looks much more similar to the one of $\Delta_{\mathbb S^2}$.

\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{figures/KhasanovaFrossardgraphonequiangularsampling.png}

	\caption{\label{fig:Equivariance error of the Frossard-Khasanove graph}Equivariance error of the Khasanova-Frossard graph on the equiangular sampling with bandwidth $b=8$ by spherical harmonic degree $\ell$.}
\end{figure}

@@

* define a measure of equiv error\\
Take a sampling scheme $V=\{v_i\in\mathbb S^2, i=0, ..., n\}$ of the sphere, a weighted undirected graph $G(V, E, \mathbf W)$, a signal $f: \mathbb S^2\to\mathbb R$ and its sampled representation $\mathbf f:\ f_i=f(v_i)$. Suppose that there exists a sampling operator $T_V: L^2(\mathbb S^2) \supset F\to \mathbb R^n,\  T_V(f) = \mathbf f$ defined on a suitable subspace $F$ of $L^2(\mathbb S^2)$ such that it is invertible, i.e., we can unambiguously reconstruct the function $f\in F$ from its sampled values $\mathbf f$. The existence of such subspace depends on the sampling scheme $V$, and its characterization is a common problem in signal processing \cite{Driscoll:1994:CFT:184069.184073}. Define the rotation operator $\Lambda(g), g\in SO(3)$: $\Lambda(g) f(\omega)=f\left(g^{-1} \omega\right)$.\\

We now want to understand how to set the edges and the weights of $G$ such that
\begin{equation}\label{eq:equivariance}
T \Lambda(g) T^{-1} \Omega_k T f = \Omega_k T \Lambda(g) f
\end{equation}
i.e., the graph convolution $\Omega_k$ and any rotation $\Lambda(g)$ commute.

Verifying equation (\ref{eq:equivariance}) is really hard in practice, due to the fact that for almost all sampling schemes $V$ it is not known if there exists a space $F$ in which $T$ is invertible. A special case is the \textit{equiangular sampling} scheme where a sampling theorem holds \cite{Driscoll:1994:CFT:184069.184073}, and thus we have a closed form of $T^{-1}$. For sampling schemes where no such sampling theorem is available, there are implementations of discrete SHT to reconstruct a sampled signal $\mathbf f$, thus providing a way to approximate $T^{-1}$. Thanks to this we are able, for a given sampling, a given function $f$, a given rotation $g$, and a given kernel $k$, to compute the \textit{normalized equivariance error} 
\begin{equation}\label{eq:equivariance error}
E_{G}(f, g) = \left(\frac{ \norm {T \Lambda(g) T^{-1} \Omega_k Tf - \Omega_k T \Lambda(g) f}_{L^2(\mathbb R^2)}}{\norm {Tf}_{L^2(\mathbb R^2)}}\right)^2
\end{equation}
where $T^{-1}$ is substituted with a discrete SHT in case $T$ is not invertible.
A measure of how equivariant a graph is with respect to rotations will then be given by the \textit{mean equivariance error}
\begin{equation}\label{eq:mean equivariance error}
\overline E_G = \mathbb E_{f, g}\ 	E_G(f, g) 
\end{equation}
In practice the expected value is obtained by averaging over a finite number of random functions and random rotations. The mean equivariance error $\overline E_G$ gives us an indication of how close the graph $G$ is from being equivariant to rotations. Now we state an intuitive concept that explains how to construct rotation invariant graphs, i.e. graphs such that $\overline{E_G}$ is small.\\
The mean equivariance error $\overline{E_G}$ will be small if the scalar product $\mathbf f^T \mathbf v_{i(\ell, m)}$ well approximates $\hat {f}(\ell,m)$ i.e., the $L^2$ scalar product of the continuous signal $\hat {f}(\ell,m)= \int_{\eta \in \mathbb S^2}f(\eta)Y_\ell^m(\eta)d\mu(\eta)$.

	\textit{If $V$ is an equal area sampling scheme}, i.e. the area around each pixel $v_i$ is the same, $\overline{E_G}$ will be small if the graph Laplacian $\mathbf L$ is such that its eigenvectors $\mathbf v_i$ well approximate the eigenfunctions of the Laplace-Beltrami operator $\Delta_{\mathbb S^2}$ evaluated in the points of the sampling scheme, i.e., 
	$$
	\mathbf v_{i(\ell, m)} \approx Y_\ell^m(x_i)
	$$

@@

* show convergence empirically (for Nside up to 1024) \\
* improved upon V1\\

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{figures/DeepSphereonHEALPix.png}	
	\includegraphics[width=\textwidth]{figures/OptimalHKGLonHEALPix.png}	
	\caption{\label{fig:DeepSphere equivariance error}Mean equivariance error of the diffusion filter $\exp(-\Lambda)$ for $G$, $G'$ and the full HKGL, by spherical harmonic degree. Notice the difference in the scale of the y axis for DeepSphere, that reaches errors up to 30\%.}
\end{figure}

* no difference in practice (see experiment xx) => NNs are resilient to equiv error\\

@@

Show: \\
* meet the desiderata \\
* DeepSphere V1 and V2 are equivalent in practice \\
* anisotropy doesn't help \\

\subsection{3D object recognition}

* same perf as other spherical CNNs, but 40 times faster \\
* compare the two samplings => not much difference \\

\subsubsection*{SHREC17}

\begin{table}[ht]
    \centering
    \begin{tabular}{l|c c c c c}
        \multicolumn{1}{l}{} & \multicolumn{2}{c}{performance} & \multicolumn{1}{c}{size} & \multicolumn{2}{c}{speed}\\
        \cmidrule(lr){2-3} \cmidrule(lr){4-4} \cmidrule(lr){5-6}
        \multicolumn{1}{l}{Method} & F1-score & mAP & params & inference [ms] & training \\ \hline
        \cite{cohen_spherical_2018} & 78.85 & 67.6 & 1.4M & 38 & 50h\\
        % Cohen \emph{s2cnn\_simple} & 78.59 & 78.85 & 400k & 12ms & 32h\\
        \cite{esteves_learning_2017} & 79.36 & 68.5 & 500k & 10 & 2h52\\ \hline
        DeepSphere \emph{Equiangular} & 79.36 & 66.5 & 190k & 1 & 2h33m \\
        DeepSphere \emph{HEALPix} & 80.65 & 68.6 & 190k & 1 & 48m\\
        DeepSphere \emph{Improved HEALPix} & ? & ? & 190k & ? & ?
    \end{tabular}
    \caption{Performance of different models for SHREC'17 task. F1-score computed with sklearn, mAP from the official script of the competition.}
    \label{tab:SHREC17_class}
\end{table}

@@

* add improved graph results too? ==> improved graph unnecessary (main information in low frequency) \\
* diminish resolution ==> no change in result until nside=8, no change for high resolution => info in the low freq\\
* same conclusion, anisotropy unnecessary cost to pay\\
* Limited by representation of data?\\

\subsubsection*{ModelNet 40}
* in appendix now\\

\subsection{cosmo}

* other sphercial CNNs cannot scale to 10M pixels (tested on 10k at most) \\
* redo the experiment with the optimal graph (cannot just copy the old results) and compare \\

@@

* small: better perf than gauge and Jiang. due to icosahedron distortion? to be confirmed\\
* cannot compare with Mudigonda (16 vs 1 input channel)\\
* full: we scale (50B pixels, 20TB) => lead to better perf?\\

\begin{table}[!ht]
\begin{tabular}{l|c c c c}
        \multicolumn{1}{l}{} & Accuracy & Average Precision & \multicolumn{2}{c}{Speed}\\
        \cmidrule(lr){2-2} \cmidrule(lr){3-3} \cmidrule(lr){4-5} 
        \multicolumn{1}{l}{Method} & mean & mean & inference [ms] & training \\ \hline
        %Mudigonda et al. & 97 & 74 & 65 & 78.67 & - & - & - \\
        \cite{jiang_spherical_2019} & 94.67 & - & - & -\\
        \cite{jiang_spherical_2019} \emph{rerun} & 94.95 & 38.41 & 10.2 & 10h14m\\
        % 328'339
        \cite{cohen_gauge_2019} (S2R) & 97.5 & 68.6 & - & -\\
        \cite{cohen_gauge_2019} (R2R) & 97.7 & 75.9 & - & -\\ \hline
        % unknown
        DeepSphere (weighted) & $97.8\pm 0.3$ & $77.15\pm 1.94$ & 25 ms & 25h \\
        DeepSphere (non-weighted) & $87.8\pm 0.5$ & $89.16\pm 1.37$ & 30 ms & 13h\\
        % 12'926'432
    \end{tabular}
    \caption{Climate pattern segmentation mean accuracy (\%), mean Average Precision (\%) and speed for different models run on the icosahedron dataset from \cite{jiang_spherical_2019}}
\end{table}

@@

\paragraph*{global regression - day in year estimation}~\\
* days represented by a sine to represent the periodic nature of the year.\\
* task too easy, temperature feature already periodic ==> second experiment with only precipitation feature\\
* structure is still helping\\

\begin{table}[!ht]
    \centering
    \begin{tabular}{c|ccc}
        order & future temp & day (temperature) & day (precipitations) \\ \hline 
        0 & 0.896 & 0.881 & - 0.920\\
        4 & 0.919 & 0.969 & 0.597\\
    \end{tabular}
    \caption{R2 coefficient for different tasks on the GHCN dataset}
    \label{tab:GHCN_results}
\end{table}

@@

* DeepSphere seems to be in the sweet spot of tradeoffs. Recall the desiderata.\\
  * respect geometry: No need for a very precise equivariance, i.e. a very good graph...\\
  * Not the most general (restricted to isotropic filters), but didn't hurt perf in any experiment\\
  * scales linearly => cannot do better ($n^{1.25}$ for best HEALPix, linear for not optimal)\\
  * flexible => independent of the sampling\\

@@

\newpage

{\LARGE \sc {Title of the paper: Supplementary Material}}
\appendix

@@

\begin{prop} 	Let $\mathcal M$ be a $k$-dimensional compact smooth manifold embedded in some euclidean space $\mathbb R^N$, and fix $p\in\mathcal M$. Let the data points $x_1, ... x_n$ be sampled form a uniform distribution on the manifold $\mathcal M$. Set $t_n=n^{-\frac{1}{k+2+\alpha}}$, for any $\alpha>0$ and let $f\in\mathcal C_\infty(\mathcal M)$. Then
	$$\frac{1}{t}\frac{1}{(4\pi t)^{k/2}} L^tf(p) \xrightarrow{t\to 0 } \frac{1}{\text{vol}(\mathcal M)}\triangle_{\mathcal M}f(p).$$
	\label{prop:3}
\end{prop}

@@

\vspace{0.5cm}
\begin{prop}\label{prop:2}
	Given a sampling regular enough i.e., for which we assume $A_i=A_j \ \forall i,j\text{ and }d^{(n)}\leq \frac{C}{\sqrt{n}}$, for a fixed $t>0$, a fixed Lipschitz function $f$ and a fixed point $y\in\mathbb S^2$ there exists a sequence $t_n = n^\beta, \beta<0$ such that 
$$
\forall f \text{ Lipschitz, } \forall x\in\mathbb S^2 \quad \left|\frac{1}{4\pi t_n^2}\left(L_n^{t_n}f(x) - L^{t_n}f(x)\right)\right|\xrightarrow{n\to \infty}0.
$$
\end{prop}
\vspace{0.5cm}

@@

The proof of this result is instructive since it shows that we need to impose some regularity conditions on the sampling. If the sampling is equal area as HEALPix, meaning that all the patches $\sigma_i$ have the same area, then we need to impose that $ d^{(n)}\leq \frac{1}{\sqrt{n}}$. If the sampling is not equal area, meaning that in general $A_i\neq A_j$, it can be shown that we need a slightly more complex condition: $\max_{i=0,...,n-1}d_iA_i\leq Cn^{-\frac{3}{2}}$.\\
In the work of Belkin et al. \cite{Belkin:2005:TTF:2138147.2138189} the sampling is drawn form a uniform random distribution on the sphere, and their proof heavily relies on the uniformity properties of the distribution from which the sampling is drawn. In our case the sampling is deterministic, and the fact that for a sphere there doesn't exist a regular sampling with more than 12 points (the vertices of a icosahedron) is indeed a problem that we need to overcome by imposing the regularity conditions above.

>>>>

%\title{Which spherical CNN should you use?\\ DeepSphere V2}
\title{DeepSphere: a graph-based spherical CNN}
%\title{DeepSphere: an efficient graph-based spherical CNN}
%\title{DeepSphere, an efficient (& equivariant) graph-based spherical CNN}

@@

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\todo}[1]{{\color[rgb]{.6,.1,.6}{#1}}}

@@

contributions:\\
* theory $\Rightarrow$ convergence (better graph)\\
* experiments on relevant problems (not spherical MNIST) $\Rightarrow$ check the desiderata\\
  * surprising: anisotropy doesn't seem useful\\

@@

\todo{
	* adjacency matrix: $A$ instead of $W$? \\
}

@@

Desiderata:\\
* respect the domain\\
  * rotational equivariance\\
  * deformation (e.g., on the icosahedron for gauge)\\
* powerful / generality $\Rightarrow$ isotropic vs anisotropic\\
* scale: computational cost \& memory usage (feature maps)\\
* flexible (samplings, irregular), simplicity (implementation)\\

\begin{table}[b]
    \centering
    \begin{tabular}{l|c|c|c|c}
         & Scale & Generality& equivariance & Flexibility \\
         \hline
        Cohen & & & &  \\
         \hline
        Gauge & & & &  \\
         \hline
        Esteves & & & &  \\
         \hline
        Jiang & & & &  \\
         \hline
        2D CNN (cubed sphere) & & & &  \\
         \hline
        Ours & & & &  \\
    \end{tabular}
    \caption{Caption}
    \label{tab:my_label}
\end{table}

On the one side, we have Cohen which is computationally very expensive but perfectly equivariant most general, on the other we have cube sphere which computationl very good but not equivariant at all. Other methods are tradedoff between these two. What is a good tradeoff? We think that DeepSphere is

Present the different spherical CNNs with pro and cons $\Rightarrow$ mostly scale vs generality\\
* 2D CNN $\Rightarrow$ doesn't respect geometry and equivariance\\
* Cohen spherical CNN $\Rightarrow$ most general, but doesn't scale\\
* Cohen gauge $\Rightarrow$ fixes scale, at the price of deformation (sphere $\Rightarrow$ icosahedron)\\
* Esteves $\Rightarrow$ ? \\
* Jiang $\Rightarrow$ need a global coordinate system (ok for planets, not projections like cosmo)\\

@@

* HEALPix: improvement from DeepSphere v1 \cite{perraudin2019deepsphere} \\
* equiangular: Renata \& Pascal

@@

\subsection{HEALPix: convergence theorem}

Given a sampling $x_0, \dots, x_{n-1}$ define $\sigma_i$ to be the patch of the surface of the sphere corresponding to the $i$-th point of the sampling, define $A_i$ to be its corresponding area and $d_i$ to be the radius of the smallest ball in $\mathbb R^3$ containing the $i$-th patch. Define $d^{(n)} := \max_{i=0, \dots, n}d_i$ and $A^{(n)}=\max_{i=0, \dots, n}A_i$.

\begin{theorem}
	For a sampling $\mathcal P = \{x_i\in\mathbb S^2\}_{i=0}^{n-1}$ of the sphere that is equi area and such that $d^{(n)} \leq \frac{C}{\sqrt{n}}$, for all $f: \mathbb S^2 \rightarrow \mathbb R$ Lipschitz with respect to the euclidean distance in $\mathbb R^3$, for all $y\in\mathbb S^2$, there exists a sequence $t_n = n^\beta$ such that the rescaled Heat Kernel Graph Laplacian operator $\frac{|\mathbb S^2|}{4\pi t_n}L^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\triangle_{\mathbb S^2}$  for $n\to\infty$:
	$$ \lim_{n\to\infty}\frac{|\mathbb S^2|}{4\pi t_n} L_n^{t_n}f(y) =  \triangle_{\mathbb S^2}f(y).$$
	\label{theo:pointwise convergence in the healpix case}
\end{theorem}

\subsection{Equiangular}

\cite{khasanova2017graphomni} designed a discrete Laplacian that is explicitly intended to work on the sphere with the equiangular sampling. They consider the set $\mathcal G$ of all the possible graphs where each node is connected only to four of its nearest neighbours (North, Sud, West, East) and propose a weighting scheme $w_{ij}$ that minimizes the difference in the response to the polynomial spectral filter $\mathcal F = \mathbf L$ evaluated on images of the same object seen at different latitudes. In other words, they solve the minimization problem
\begin{equation} \label{eq:minimization frossard}
	\min_{W\in\mathcal G} \left|\mathcal{F}\left(\mathbf{y}\left(v_{ e}\right)\right)-\mathcal{F}\left(\mathbf{y}\left(v_{ i}\right)\right)\right|
\end{equation}
for the adjacency matrix $W$, where $\mathbf y(v_i)$ is the image of the object on the sphere centered on the vertex $v_i$, and $\mathcal F (\mathbf y(v_e))$ is the response of the filter at the vertex $v_e$ that lies at the same longitude of the vertex $v_i$ but on the equator. In their work they prove that the optimal weights solving the minimization problem (\ref{eq:minimization frossard}) are given by weights $w_{ij}$ inversely proportional to the Euclidean distance between vertices:
\begin{equation} \label{eq:frossard weights}
	w_{ij} = \frac{1}{\norm{x_i-x_j}}
\end{equation}
This construction is interesting since it is adapted to the equiangular sampling, and leads to a very sparse graph with only 4 neighbors per vertex. Furthermore, to obtain the weights (\ref{eq:frossard weights}) every calculation was done in the \textit{spatial domain}, without any consideration about the spectral interpretation of the filter. In order to compare it to the HKGL we show the equivariance error by spherical harmonic degree in figure \ref{fig:Equivariance error of the Frossard-Khasanove graph}. It can be appreciated how this construction performs a little worse that the full HKGL for low bandwidth samplings, but converges faster than the \textit{full} HKGL, and its spectrum looks much more similar to the one of $\Delta_{\mathbb S^2}$.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{figures/KhasanovaFrossardgraphonequiangularsampling.png}
	\caption{Equivariance error of the Khasanova-Frossard graph on the equiangular sampling with bandwidth $b=8$ by spherical harmonic degree $\ell$.}
	\label{fig:Equivariance error of the Frossard-Khasanove graph}
\end{figure}

@@

\todo{
* define a measure of equiv error \\
* show convergence empirically (for Nside up to 1024) \\
* improved upon V1 \\
* no difference in practice (see experiment xx) $\Rightarrow$ NNs are resilient to equiv error \\
}

Take a sampling scheme $V=\{v_i\in\mathbb S^2, i=0, \dots, n\}$ of the sphere, a weighted undirected graph $G(V, E, \mathbf W)$, a signal $f: \mathbb S^2\to\mathbb R$ and its sampled representation $\mathbf f:\ f_i=f(v_i)$. Suppose that there exists a sampling operator $T_V: L^2(\mathbb S^2) \supset F\to \mathbb R^n,\  T_V(f) = \mathbf f$ defined on a suitable subspace $F$ of $L^2(\mathbb S^2)$ such that it is invertible, i.e., we can unambiguously reconstruct the function $f\in F$ from its sampled values $\mathbf f$. The existence of such subspace depends on the sampling scheme $V$, and its characterization is a common problem in signal processing \cite{Driscoll:1994:CFT:184069.184073}. Define the rotation operator $\Lambda(g), g\in SO(3)$: $\Lambda(g) f(\omega) = f\left(g^{-1} \omega\right)$.

We now want to understand how to set the edges and the weights of $G$ such that
\begin{equation} \label{eq:equivariance}
	T \Lambda(g) T^{-1} \Omega_k T f = \Omega_k T \Lambda(g) f,
\end{equation}
i.e., the graph convolution $\Omega_k$ and any rotation $\Lambda(g)$ commute.

Verifying equation (\ref{eq:equivariance}) is really hard in practice, due to the fact that for almost all sampling schemes $V$ it is not known if there exists a space $F$ in which $T$ is invertible. A special case is the \textit{equiangular sampling} scheme where a sampling theorem holds \cite{Driscoll:1994:CFT:184069.184073}, and thus we have a closed form of $T^{-1}$. For sampling schemes where no such sampling theorem is available, there are implementations of discrete SHT to reconstruct a sampled signal $\mathbf f$, thus providing a way to approximate $T^{-1}$. Thanks to this we are able, for a given sampling, a given function $f$, a given rotation $g$, and a given kernel $k$, to compute the \textit{normalized equivariance error}
\begin{equation} \label{eq:equivariance error}
	E_{G}(f, g) = \left(\frac{ \norm {T \Lambda(g) T^{-1} \Omega_k Tf - \Omega_k T \Lambda(g) f}_{L^2(\mathbb R^2)}}{\norm {Tf}_{L^2(\mathbb R^2)}}\right)^2,
\end{equation}
where $T^{-1}$ is substituted with a discrete SHT in case $T$ is not invertible.
A measure of how equivariant a graph is with respect to rotations will then be given by the \textit{mean equivariance error}
\begin{equation} \label{eq:mean equivariance error}
	\overline E_G = \mathbb E_{f, g} \ E_G(f, g).
\end{equation}
In practice the expected value is obtained by averaging over a finite number of random functions and random rotations. The mean equivariance error $\overline E_G$ gives us an indication of how close the graph $G$ is from being equivariant to rotations. Now we state an intuitive concept that explains how to construct rotation invariant graphs, i.e., graphs such that $\overline{E_G}$ is small.
The mean equivariance error $\overline{E_G}$ will be small if the scalar product $\mathbf f^T \mathbf v_{i(\ell, m)}$ well approximates $\hat {f}(\ell,m)$, i.e., the $L^2$ scalar product of the continuous signal $\hat {f}(\ell,m)= \int_{\eta \in \mathbb S^2}f(\eta)Y_\ell^m(\eta)d\mu(\eta)$.

\textit{If $V$ is an equal area sampling scheme}, i.e., the area around each pixel $v_i$ is the same, $\overline{E_G}$ will be small if the graph Laplacian $\mathbf L$ is such that its eigenvectors $\mathbf v_i$ well approximate the eigenfunctions of the Laplace-Beltrami operator $\Delta_{\mathbb S^2}$ evaluated in the points of the sampling scheme, i.e., $\mathbf v_{i(\ell, m)} \approx Y_\ell^m(x_i)$.

@@

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{figures/DeepSphereonHEALPix.png}
	\includegraphics[width=\textwidth]{figures/OptimalHKGLonHEALPix.png}
	\caption{\label{fig:DeepSphere equivariance error}Mean equivariance error of the diffusion filter $\exp(-\Lambda)$ for $G$, $G'$ and the full HKGL, by spherical harmonic degree. Notice the difference in the scale of the y axis for DeepSphere, that reaches errors up to 30\%.}
\end{figure}

@@

\todo{
Show: \\
* meet the desiderata \\
* DeepSphere V1 and V2 are equivalent in practice, wait on SHREC and cosmo exp \\
* anisotropy doesn't help \\
}

\subsection{3D objects recognition}

\todo{
* same perf as other spherical CNNs, but 40 times faster \\
* compare samplings (equi, HEALPix) $\Rightarrow$ not much difference \\
* compare with improved graph (Nside=32) $\Rightarrow$ no improvement if information in low frequencies, do an SHT \\
* ModelNet 40: same conclusions $\Rightarrow$ appendix \\
}

\begin{table}
    \centering
    \begin{tabular}{l|c c c c c}
        \multicolumn{1}{l}{} & \multicolumn{2}{c}{performance} & \multicolumn{1}{c}{size} & \multicolumn{2}{c}{speed}\\
        \cmidrule(lr){2-3} \cmidrule(lr){4-4} \cmidrule(lr){5-6}
        \multicolumn{1}{l}{Method} & F1 & mAP & params & inference [ms] & training \\ \hline
        \cite{cohen_spherical_2018} & 78.85 & 67.6 & 1.4M & 38 & 50h\\
        % Cohen \emph{s2cnn\_simple} & 78.59 & 78.85 & 400k & 12ms & 32h\\
        \cite{esteves_learning_2017} & 79.36 & 68.5 & 500k & 10 & 2h52\\ \hline
        DeepSphere \emph{Equiangular} & 79.36 & 66.5 & 190k & 1 & 2h33m \\
        DeepSphere \emph{HEALPix} & 80.65 & 68.6 & 190k & 1 & 48m\\
        DeepSphere \emph{Improved HEALPix} & ? & ? & 190k & ? & ?
    \end{tabular}
    \caption{Performance of different models for SHREC'17 task. F1-score computed with sklearn, mAP from the official script of the competition.}
    \label{tab:SHREC17_class}
\end{table}

@@

* diminish resolution $\Rightarrow$ no change in result until nside=8, no change for high resolution $\Rightarrow$ info in the low freq\\
* same conclusion, anisotropy unnecessary cost to pay\\
* Limited by representation of data? yep, treat them as point clouds \\

\subsection{Cosmological model discrimination}

\todo{
* other sphercial CNNs cannot scale to 10M pixels (tested on 10k at most) \\
* redo the experiment with the optimal graph \\
}

@@

* small: better perf than gauge and Jiang. due to icosahedron distortion? to be confirmed\\
* cannot compare with Mudigonda (16 vs 1 input channel)\\
* (full: we scale (50B pixels, 20TB) $\Rightarrow$ lead to better perf?) \\

\begin{table}
	\begin{tabular}{l|c c c c}
	\multicolumn{1}{l}{} & Accuracy & Average Precision & \multicolumn{2}{c}{Speed}\\
	\cmidrule(lr){2-2} \cmidrule(lr){3-3} \cmidrule(lr){4-5}
	\multicolumn{1}{l}{Method} & mean & mean & inference [ms] & training \\ \hline
	%Mudigonda et al. & 97 & 74 & 65 & 78.67 & - & - & - \\
	\cite{jiang_spherical_2019} & 94.67 & - & - & -\\
	\cite{jiang_spherical_2019} \emph{rerun} & 94.95 & 38.41 & 10.2 & 10h14m\\
	% 328'339
	\cite{cohen_gauge_2019} (S2R) & 97.5 & 68.6 & - & -\\
	\cite{cohen_gauge_2019} (R2R) & 97.7 & 75.9 & - & -\\ \hline
	% unknown
	DeepSphere (weighted) & $97.8\pm 0.3$ & $77.15\pm 1.94$ & 25 ms & 25h \\
	DeepSphere (non-weighted) & $87.8\pm 0.5$ & $89.16\pm 1.37$ & 30 ms & 13h\\
	% 12'926'432
	\end{tabular}
	\caption{Climate pattern segmentation mean accuracy (\%), mean Average Precision (\%) and speed for different models run on the icosahedron dataset from \cite{jiang_spherical_2019}}
\end{table}

@@

\paragraph*{global regression - day in year estimation}~\\
* days represented by a sine to represent the periodic nature of the year.\\
* task too easy, temperature feature already periodic $\Rightarrow$ second experiment with only precipitation feature\\
* structure is still helping\\

\begin{table}
    \centering
    \begin{tabular}{c|ccc}
        order & future temp & day (temperature) & day (precipitations) \\ \hline
        0 & 0.896 & 0.881 & - 0.920\\
        4 & 0.919 & 0.969 & 0.597\\
    \end{tabular}
    \caption{R2 coefficient for different tasks on the GHCN dataset}
    \label{tab:GHCN_results}
\end{table}

@@

* DeepSphere seems to be in the sweet spot of tradeoffs. Recall the desiderata.\\
  * respect geometry: No need for a very precise equivariance, i.e., a very good graph...\\
  * Not the most general (restricted to isotropic filters), but didn't hurt perf in any experiment\\
  * scales linearly $\Rightarrow$ cannot do better ($n^{1.25}$ for best HEALPix, linear for not optimal)\\
  * flexible $\Rightarrow$ independent of the sampling\\

@@

\newpage
{\LARGE \sc {Supplementary Material}}
\appendix

@@

\begin{prop} 	Let $\mathcal M$ be a $k$-dimensional compact smooth manifold embedded in some euclidean space $\mathbb R^N$, and fix $p\in\mathcal M$. Let the data points $x_1, \dots, x_n$ be sampled form a uniform distribution on the manifold $\mathcal M$. Set $t_n=n^{-\frac{1}{k+2+\alpha}}$, for any $\alpha>0$ and let $f\in\mathcal C_\infty(\mathcal M)$. Then
	$$\frac{1}{t}\frac{1}{(4\pi t)^{k/2}} L^tf(p) \xrightarrow{t\to 0 } \frac{1}{\text{vol}(\mathcal M)}\triangle_{\mathcal M}f(p).$$
	\label{prop:3}
\end{prop}

@@

\vspace{0.5cm}
\begin{prop}\label{prop:2}
	Given a sampling regular enough, i.e., for which we assume $A_i=A_j \ \forall i,j\text{ and }d^{(n)}\leq \frac{C}{\sqrt{n}}$, for a fixed $t>0$, a fixed Lipschitz function $f$ and a fixed point $y\in\mathbb S^2$ there exists a sequence $t_n = n^\beta, \beta<0$ such that
$$
\forall f \text{ Lipschitz, } \forall x\in\mathbb S^2 \quad \left|\frac{1}{4\pi t_n^2}\left(L_n^{t_n}f(x) - L^{t_n}f(x)\right)\right|\xrightarrow{n\to \infty}0.
$$
\end{prop}
\vspace{0.5cm}

@@

The proof of this result is instructive since it shows that we need to impose some regularity conditions on the sampling. If the sampling is equal area as HEALPix, meaning that all the patches $\sigma_i$ have the same area, then we need to impose that $ d^{(n)}\leq \frac{1}{\sqrt{n}}$. If the sampling is not equal area, meaning that in general $A_i\neq A_j$, it can be shown that we need a slightly more complex condition: $\max_{i=0,\dots,n-1}d_iA_i\leq Cn^{-\frac{3}{2}}$.\\
In the work of Belkin et al. \cite{Belkin:2005:TTF:2138147.2138189} the sampling is drawn form a uniform random distribution on the sphere, and their proof heavily relies on the uniformity properties of the distribution from which the sampling is drawn. In our case the sampling is deterministic, and the fact that for a sphere there doesn't exist a regular sampling with more than 12 points (the vertices of a icosahedron) is indeed a problem that we need to overcome by imposing the regularity conditions above.
